Palomar 5 Gaps

### 2024-09-19 paper/main.tex
small edits, replacing XXXs

  In this study, we aim to investigate the potential influence of globular clusters on stellar streams, specifically examining whether galactic globular clusters can produce gaps similar to those expected from dark matter subhalos. If globular clusters do create significant gaps, this must be considered when using stellar streams to infer the local presence of dark matter. Our focus is on the Palomar 5 stream, which has been well-studied and has observable gaps, a high membership count, and favorable visibility due to its position above the galactic disk. In the following sections, we describe the modeling of this stream, the gravitational interactions with the galaxy and other globular clusters, and the statistical analysis of the perturbations affecting the stream.

@@

    To generate initial conditions, we treat the five-dimensional phase space variables (position and velocity components) for each globular cluster as Gaussian distributions. Using the \texttt{np.random.gauss} function, we generate 50 samples per cluster, producing 50 sets of initial conditions for the entire globular cluster system. These samples are stored in data files for later use.

@@

    For the galactic potential, we employed the static model proposed by Pouliasis et al. (2017), which consists of a superposition of a thin disk, thick disk, and a dark matter halo. This model is assumed to be time-independent over the course of our simulations.

    The primary methodological departure from Ferrone et al. (2023) is that now we include globular cluster interactions. Instead of treating the clusters as point masses that only experience the gravitational potential of the galaxy, we now compute N-body interactions between the clusters during the backward integration over 5 billion years. The equation of motion for the globular clusters is thus: 
    \begin{equation}
      \ddot{\vec{r}}_i = -\nabla \Phi(R_i,z_i) + \left.\sum_{j\neq i} \frac{Gm_j}{\left(|\vec{r}_j - \vec{r}_i|^2 + b_j^2\right)^{3/2}}\right. \left(\vec{r}_j - \vec{r}_i\right).
      \end{equation}
    Although these interactions were included, we observed no significant exchanges of orbital energy or momentum between clusters. The resulting orbits were saved to data files for subsequent analysis.

    At the computed positions 5 Gyr ago, each globular cluster was replaced by a system of 100,000 particles, sampled from a Plummer distribution to represent the stellar content of the cluster. We then integrated the evolution of these particles forward in time to the present day. During this forward integration, each star particle feels the gravitational influence of the galaxy and all globular clusters, but not other star particles. We do not perform direct N-body simulations between star particles; instead, at each time step, we compute the positions of all globular clusters and apply their forces to the star particles. The equation of motion is thus: 
    \begin{equation}
      \ddot{\vec{r}}_p = -\nabla \Phi(R_p,z_p) + \left.\sum_{j} \frac{Gm_j}{\left(|\vec{r}_j(t) - \vec{r}_p|^2 + b_j^2\right)^{3/2}}\right. \left(\vec{r}_j(t)- \vec{r}_p\right)
      \end{equation}

@@

  \begin{figure*}
    \centering
    \includegraphics[width=\linewidth]{stream_on_sky_Pal5_monte-carlo-009_pouliasis2017pii-GCNBody.png}
    \caption[]{A model of the Palomar 5 stream in an axis-symmetric galactic potential plus all other globular clusters. The top plot is a scatter plot of star-particles that escaped the cluster due to tidal forces, the bottom plot shows the marginalized 1D density profile in longitude. Two large gaps are present and are due to the passage of two globular clusters. }
    \end{figure*}

>>>>

  In this study, we aim to investigate the potential influence of globular clusters on stellar streams, specifically examining whether galactic globular clusters can produce gaps similar to those expected from dark matter subhalos. If globular clusters do create significant gaps, this must be considered when using stellar streams to infer the local presence of dark matter. Our focus is on the Palomar 5 stream, which has been well-studied and has observable gaps, a high membership count, and favorable visibility due to its position above the galactic disk. We show a result from one our simulations in \ref{fig:stream_on_sky}. In the following sections, we describe the modeling of this stream, the gravitational interactions with the galaxy and other globular clusters, and the statistical analysis of the perturbations affecting the stream.

  \begin{figure*}
    \centering
    \includegraphics[width=\linewidth]{stream_on_sky_Pal5_monte-carlo-009_pouliasis2017pii-GCNBody.png}
    \caption[]{A model of the Palomar 5 stream in an axis-symmetric galactic potential plus all other globular clusters. The top plot is a scatter plot of star-particles that escaped the cluster due to tidal forces, the bottom plot shows the marginalized 1D density profile in longitude. Two large gaps are present and are due to the passage of two globular clusters. }
    \label{fig:stream_on_sky}
    \end{figure*}

@@

    To generate initial conditions, we treat the four phase space variables with uncertainties (distance, line-of-sight velocities, and proper motions) and mass for each globular cluster as five-dimensional Gaussian distributions. Using the \texttt{np.random.gauss} function, we generate 50 samples per cluster, producing 50 sets of initial conditions for the entire globular cluster system. These samples are stored in data files for later use.

@@

    For the galactic potential, we employed the second static model proposed by Pouliasis et al. (2017), which consists of a superposition of a thin disk, thick disk, and a dark matter halo. This model is assumed to be time-independent over the course of our simulations.

    The primary methodological departure from Ferrone et al. (2023) is that now we include globular cluster interactions. Instead of treating the clusters as point masses that only experience the gravitational potential of the galaxy, we now compute N-body interactions between the clusters during the backward integration over 5 billion years. The equation of motion for the globular clusters is thus: 
    \begin{equation}
      \ddot{\vec{r}}_i = -\nabla \Phi(R_i,z_i) + \left.\sum_{j\neq i} \frac{Gm_j}{\left(|\vec{r}_j - \vec{r}_i|^2 + b_j^2\right)^{3/2}}\right. \left(\vec{r}_j - \vec{r}_i\right).
      \end{equation}\label{eq:GCNBody}
    Although these interactions were included, we observed no significant exchanges of orbital energy or momentum between clusters. The resulting orbits were saved to data files for subsequent analysis. We integrate clusters into the past by 5 Gyr.

    At the computed positions 5 Gyr ago, each globular cluster was gifted a system of 100,000 particles, sampled from a Plummer distribution to represent the stellar content of the cluster. We then integrated the evolution of these particles forward in time to the present day. During this forward integration, each star particle feels the gravitational influence of the galaxy and all globular clusters, but not other star particles. We do not perform direct N-body simulations between star particles as in eq.~\ref{eq:GCNBody}; instead, at each time step, we compute the positions of all globular clusters and apply their forces to the star particles. The equation of motion is thus: 
    \begin{equation}
      \ddot{\vec{r}}_p = -\nabla \Phi(R_p,z_p) + \left.\sum_{j} \frac{Gm_j}{\left(|\vec{r}_j(t) - \vec{r}_p|^2 + b_j^2\right)^{3/2}}\right. \left(\vec{r}_j(t)- \vec{r}_p\right)
      \end{equation}

@@
